---
title: "Extended Methods for 'Status Invisibility Improves Emotional Well-being and Alleviates its Economic Gradient in Social Network Experiments'"
author: "Akihiro Nishi, Christopher A. German, Sage K. Iwamoto, and Nicholas A. Christakis"
date: "11/7/2022"
output: pdf_document
---

## 1. Install and load some packages  
We have installed packages below (that are going to be used.)
```{r}
#install.packages("lme4")
library(lme4) #multi-level modeling
library(lmerTest) #testing in multi-level modeling
library(mediation) #causal mediation analysis
```

## 2. Data setting

### 2.1. Import the necessary data

The two .csv files for replicating the main results are available on AN’s Github page (https://github.com/akihironishi).
```{r}
setwd("/Users/akihironishi/Dropbox/ArticlesAN/UR-AN14HAPP-ProcB/NEW/Data/")
#Change the directory by yourself
HappStandard = read.csv("HappStandard.csv")
#HappStandard for the first series of experiments (no harming option)
HarmStandard = read.csv("HarmStandard.csv")
#HarmStandard for the second series of experiments (harming option)
```

### 2.2. The list of variables  

**game**: game ID (top digit is 1: first series, 2: second series)  
**superid**: individual ID (top three digits are game ID)  
**round**: round 0-15  
**showScore**: 0 for invisible and 1 for visible  
**initScore**: 200 for the poor and 1150 for the rich  
**payoff**: payoff at each round  
**cumulativePayoff**: cumulative payoff (absolute wealth)  
**cPayoffS**: wealth standardized (relative wealth)
**WealthLevel**: wealth category (5 caegories based on SD)  
**coop**: 1 for cooperating  
**local_rate_coop**: the proportion of cooperation among the connecting neighbors of each study participant  
**degree**: The number of social ties (connecting neighbors)  
**happ**: Subjective Well-being (-2 to 2)  
**harm**: 1 for harming  

## 3. Main regression analysis  

### 3.1. First series (with and without an interaction term): HappStandard   

```{r}
#invisible only
summary(lmer(data = HappStandard[HappStandard$showScore==0,], happ ~ cPayoffS
             + factor(round) + (1|game) + (1|superid)))$coef
#visible only
summary(lmer(data = HappStandard[HappStandard$showScore==1,], happ ~ cPayoffS
             + factor(round) + (1|game) + (1|superid)))$coef
#invisible + visible, without an interaction term
summary(lmer(data = HappStandard, happ ~ showScore + cPayoffS + factor(round)
             + (1|game) + (1|superid)))$coef
#invisible + visible, with an interaction term
summary(lmer(data = HappStandard, happ ~ showScore * cPayoffS + factor(round)
             + (1|game) + (1|superid)))$coef
```

### 3.2. Second series (with and without an interaction term): HarmStandard  

```{r}
#invisible only
summary(lmer(data = HarmStandard[HarmStandard$showScore==0,], happ ~ cPayoffS
             + factor(round) + (1|game) + (1|superid)))$coef
#visible only
summary(lmer(data = HarmStandard[HarmStandard$showScore==1,], happ ~ cPayoffS
             + factor(round) + (1|game) + (1|superid)))$coef
#invisible + visible, without an interaction term
summary(lmer(data = HarmStandard, happ ~ showScore + cPayoffS + factor(round)
             + (1|game) + (1|superid)))$coef
#invisible + visible, with an interaction term
summary(lmer(data = HarmStandard, happ ~ showScore * cPayoffS + factor(round)
             + (1|game) + (1|superid)))$coef
```

## 4. Causal mediation analysis  

We use the "local_rate_coop" variable as a potential mediator of the showScore - happ association.  

```{r}
detach(package:lmerTest, unload=TRUE)
#COMMENT: lmerTest package has a conflict with mediation package
```

### 4.1. HappStandard: First series (with different clusters)    

```{r}
happ.mediator.fit <- lmer(data = HappStandard, local_rate_coop ~ showScore 
                          + cPayoffS + factor(round) + (1|superid))
happ.outcome.fit <- lmer(data = HappStandard, happ ~ local_rate_coop 
                         + showScore + cPayoffS + factor(round) + (1|superid))
happ_med <- mediate(happ.mediator.fit, happ.outcome.fit, 
                    treat = "showScore", mediator = "local_rate_coop")
#"mediate does not support more than two levels per model" 
#-> a lower cluster was chosen.
summary(happ_med) #Prop. Mediated: 0.1705 (output will vary due to simulations)

#REPRODUCTION with a different cluster
happ.mediator.fit <- lmer(data = HappStandard, local_rate_coop ~ showScore 
                          + cPayoffS + factor(round) + (1|game))
happ.outcome.fit <- lmer(data = HappStandard, happ ~ local_rate_coop 
                         + showScore + cPayoffS + factor(round) + (1|game))
happ_med <- mediate(happ.mediator.fit, happ.outcome.fit, 
                    treat = "showScore", mediator = "local_rate_coop")
summary(happ_med) #Prop. Mediated: 0.12536 (a little bit attenuated)
```

### 4.2. HarmStandard: Second series (with different clusters)  

```{r}
harm.mediator.fit <- lmer(data = HarmStandard, local_rate_coop ~ showScore 
                          + cPayoffS + factor(round) + (1|superid))
harm.outcome.fit <- lmer(data = HarmStandard, happ ~ local_rate_coop 
                         + showScore +  cPayoffS + factor(round) + (1|superid))
harm_med <- mediate(harm.mediator.fit, harm.outcome.fit, 
                    treat = "showScore", mediator = "local_rate_coop")
summary(harm_med) #Prop. Mediated: 0.3744 (output will vary due to simulations)

#REPRODUCTION with a different cluster
harm.mediator.fit <- lmer(data = HarmStandard, local_rate_coop ~ showScore 
                          + cPayoffS + factor(round) + (1|game))
harm.outcome.fit <- lmer(data = HarmStandard, happ ~ local_rate_coop 
                         + showScore + cPayoffS + factor(round) + (1|game))
harm_med <- mediate(harm.mediator.fit, harm.outcome.fit, 
                    treat = "showScore", mediator = "local_rate_coop")
summary(harm_med) #Prop. Mediated: 0.251 (a little bit attenuated)
```

```{r}
library(lmerTest) #make lmerTest package back for further analysis
```

## 5. Additional analysis  

### 5.1. Cooperation rate over the 15 rounds (a dichotomous variable)  

```{r}
#First series
happ_data1 = HappStandard[HappStandard$round>=1,
                          c("game","superid","showScore","round","coop")]
happ_data2 = reshape(happ_data1,timevar="round",
                     idvar=c("game","superid","showScore"),direction="wide")
happ_data2$coop_rate = apply(happ_data2[,substr(names(happ_data2),1,4)=="coop"],
                             1,function(x) mean(x,na.rm=T))
summary(lmer(data = happ_data2, coop_rate ~ showScore + (1|game)))$coef

#Second series
harm_data1 = HarmStandard[HarmStandard$round>=1,
                          c("game","superid","showScore","round","coop")]
harm_data2 = reshape(as.data.frame(harm_data1),timevar="round",
                     idvar=c("game","superid","showScore"),direction="wide")
harm_data2$coop_rate = apply(harm_data2[,substr(names(harm_data2),1,4)=="coop"],
                             1,function(x) mean(x,na.rm=T))
summary(lmer(data = harm_data2, coop_rate ~ showScore + (1|game)))$coef
```

### 5.2. Happiness comparison (first series v.s. second series)  

```{r}
HappStandard$series = "happ"
HarmStandard$series = "harm"
totaldata = rbind(HappStandard,HarmStandard)
summary(lmer(data = totaldata, happ ~ series + factor(round)
             + (1|game) + (1|superid)))$coef
```

### 5.3. Harming option execution rate over the 15 rounds  

```{r}
harm_data1H = HarmStandard[HarmStandard$round>=1,
                           c("game","superid","showScore","round","harm")]
harm_data2H = reshape(as.data.frame(harm_data1H),timevar="round",
                      idvar=c("game","superid","showScore"),direction="wide")
harm_data2H$harm_rate = apply(harm_data2H[,substr(names(harm_data2H),1,4)=="harm"]
                              ,1,function(x) mean(x,na.rm=T))
summary(lmer(data = harm_data2H, harm_rate ~ showScore + (1|game)))$coef
```

### 5.4. Follow-up analysis for Easterlin's Paradox  

**Question 1**: Did happiness increase as the rounds progressed in general?   
**A**: No.  

```{r}
summary(lmer(data = HappStandard, happ ~ showScore + cPayoffS + round 
             + (1|game) + (1|superid)))$coef
summary(lmer(data = HarmStandard, happ ~ showScore + cPayoffS + round
             + (1|game) + (1|superid)))$coef
```

**Question 2**: After relative wealth is incorporated into the model, did absolute wealth matter?   
**A**: No, relative wealth has little predictive ability on the outcome.  

```{r}
summary(lmer(data = HappStandard, happ ~ showScore + cPayoffS 
             + cumulativePayoff + round + (1|game) + (1|superid)))$coef
summary(lmer(data = HarmStandard, happ ~ showScore + cPayoffS 
             + cumulativePayoff + round + (1|game) + (1|superid)))$coef
#Both Ps for cumulativePayoff (absolute wealth) > 0.10
```